Automated verification of stand-alone solar photovoltaic systems

### 2018-11-21 Tex/LaTeX2e Style File/Trindade_IEEE_journal_r2.tex
paper update

%--------------------------------------
\subsection{ESBMC}
%--------------------------------------

ESBMC (or Efficient SMT-based Bounded Model Checker) is an open source, permissively licensed (Apache 2), cross platform bounded model checking for C and C++ programs \cite{esbmc2018}. 
ESBMC's verification flow can be summarized in three stages: (I) a front-end that can read and compile C/C++ code, where the formal specification of the system to be verified is first treated; (II) preprocessing steps to deal with the representation of the code, control flow and unwinding of loops, and the simplification of the algorithm, aiming to reduce the computational complexity; and finally (III) the solver, where all the constraints and properties of the system to be verified are encoded into SMT and checked for satisfiability.
%%The efficient SMT based model checker, is a software verification tool for C and C++ code bases. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Fig. \ref{fig:esbmcarch} shows the tool architecture. 
%\includegraphics[width=0.4\textwidth]{ESBMCarch4}
%\caption{ESBMC architecture (adapted from \cite{esbmc2018}).}
%\label{fig:esbmcarch}
%ESBMC uses clang as front-end, a state-of-the-art compiler suite for C/C++/ObjectiveC/ObjectiveC++ widely used in industry \cite{esbmc2018}. % and generate an Abstract Syntax Tree (AST). %%%%%One legacy CBMC-based front-end that supports both C and C++ and a new clang-based front-end that currently only supports C. The data types and bit vectors are created in the front-end when parsing the code. 
%%%%%%%Regardless of the chosen front-end, 
%The output is an AST (tree representation of the abstract syntactic structure of source code written in a programming language) that will be used by the GOTO converter to generate a GOTO program, which has simplified control flow and is suitable for bounded unwinding. The next step is the symbolic execution, when the GOTO program is executed (unrolling loops up to the bound $ k $) and converted to Static Single Assignments (SSA) form. SSA form enables the simplification of algorithms and reduction of computational complexity. %%%%%%%%%, where each variable is a target of exactly one assignment statement in the program text. 
%The feature called k-induction at Fig. \ref{fig:esbmcarch} allows the model checker to find a property violation or even to prove (partial) correctness without fully unwinding loops. 
%%
%%%%%%%During the symbolic execution, ESBMC aggressively tries to simplify the program \cite{Ramalho}; it propagates all constants and solves any assertions that can be statically determined. This is an important step for the verification, since ESBMC can fully verify programs without calling a solver, if the inputs are deterministic. That characteristic is useful at this work to perform the sizing check of the PV system project. 
%Following, the SSA expressions are then encoded using the chosen SMT solver supported by ESBMC. At this paper, it was used the Z3 solver \cite{DeMoura}.
%%%%%: Boolector (default) \cite{Brummayer}, Z3 \cite{DeMoura}, MathSAT \cite{Cimatti}, Yices \cite{Dutertre} and CVC4 \cite{Barrett}. 
%Finally, the system attempting to determine whether a formula, which is the disjunction of all possible errors, can be satisfied. 
If the SMT formula is shown to be satisfiable (SAT), a counterexample is presented; otherwise, the formula is unsatisfiable (UNSAT), i.e., there are no errors up to the given unwinding bound. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5%
%%%%%ESBMC can be invoked through the command-line interface or configured through the Eclipse plug-in. Specific problems and the solver can be selected that way. 
%ESBMC aims to support all of ISO/IEC 9899:2017 C programming language standard, and detects errors in software by simulating a finite prefix of the program execution with all possible inputs. Classes of problems that can be detected include: user specified assertion failures, out of bounds array access, illegal pointer dereferences, double-free of malloc'd memory, misaligned memory access; integer overflows, divide by zero, memory leaks, concurrent software, deadlock, and data races (i.e. competing writes).

>>>>

%--------------------------------------
\subsection{ESBMC}
%--------------------------------------
ESBMC (or Efficient SMT-based Bounded Model Checker) is an open source, permissively licensed (Apache 2), cross platform bounded model checking for C and C++ programs \cite{esbmc2018}. 
%
ESBMC's verification flow can be summarized in three stages: (I) a front-end that can read and compile C/C++ code, where the formal specification of the system to be verified is first treated; (II) preprocessing steps to deal with the representation of the code, control flow and unwinding of loops, and the simplification of the algorithm, aiming to reduce the computational complexity; and finally (III) the solver, or SMT solver, where all the constraints and properties of the system to be verified are encoded into SMT and checked for satisfiability.
%
%%The efficient SMT based model checker, is a software verification tool for C and C++ code bases. 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Fig. \ref{fig:esbmcarch} shows the tool architecture. 
%
%\begin{figure}[h]
%\includegraphics[width=0.4\textwidth]{ESBMCarch4}
%\centering
%\caption{ESBMC architecture (adapted from \cite{esbmc2018}).}
%\label{fig:esbmcarch}
%\end{figure}
%ESBMC uses clang as front-end, a state-of-the-art compiler suite for C/C++/ObjectiveC/ObjectiveC++ widely used in industry \cite{esbmc2018}. % and generate an Abstract Syntax Tree (AST). %%%%%One legacy CBMC-based front-end that supports both C and C++ and a new clang-based front-end that currently only supports C. The data types and bit vectors are created in the front-end when parsing the code. 
%%%%%%%Regardless of the chosen front-end, 
%The output is an AST (tree representation of the abstract syntactic structure of source code written in a programming language) that will be used by the GOTO converter to generate a GOTO program, which has simplified control flow and is suitable for bounded unwinding. The next step is the symbolic execution, when the GOTO program is executed (unrolling loops up to the bound $ k $) and converted to Static Single Assignments (SSA) form. SSA form enables the simplification of algorithms and reduction of computational complexity. %%%%%%%%%, where each variable is a target of exactly one assignment statement in the program text. 
%The feature called k-induction at Fig. \ref{fig:esbmcarch} allows the model checker to find a property violation or even to prove (partial) correctness without fully unwinding loops. 
%%
%%%%%%%During the symbolic execution, ESBMC aggressively tries to simplify the program \cite{Ramalho}; it propagates all constants and solves any assertions that can be statically determined. This is an important step for the verification, since ESBMC can fully verify programs without calling a solver, if the inputs are deterministic. That characteristic is useful at this work to perform the sizing check of the PV system project. 
%Following, the SSA expressions are then encoded using the chosen SMT solver supported by ESBMC. At this paper, it was used the Z3 solver \cite{DeMoura}.
%%%%%: Boolector (default) \cite{Brummayer}, Z3 \cite{DeMoura}, MathSAT \cite{Cimatti}, Yices \cite{Dutertre} and CVC4 \cite{Barrett}. 
%
%Finally, the system attempting to determine whether a formula, which is the disjunction of all possible errors, can be satisfied. 
If the SMT formula is shown to be satisfiable (SAT), a counterexample is presented; otherwise, the formula is unsatisfiable (UNSAT), i.e., there are no errors up to the given unwinding bound.

% incremental ESBMC text here
Related to the standardized input language to SMT solvers, aka SMTLIB logic format, there is a resource called \textit{assertion stack}. An assertion, in SMT solvers, is a constraint on the variables in formula that must hold if the formula is satisfiable \cite{Morse2015}. New assertions can be added to or old assertions removed from this stack, depending on the evaluated value of variables. This enables the solver to learn from previous checks for future checks, optimizing the search procedure and potentially eliminating a large amount of formula state space to be searched, because it solves and discart data during the process, incrementally. This technique is called "Incremental SMT" \cite{Morse2015}. The process of retraction to build and deconstruct the formula allow reduced SMT-conversion overhead of memory, mainly when the verified system is complex and the computing platform do not have large amount of RAM memory to deal with all the design space state.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5%
%%%%%ESBMC can be invoked through the command-line interface or configured through the Eclipse plug-in. Specific problems and the solver can be selected that way. 
%
%ESBMC aims to support all of ISO/IEC 9899:2017 C programming language standard, and detects errors in software by simulating a finite prefix of the program execution with all possible inputs. Classes of problems that can be detected include: user specified assertion failures, out of bounds array access, illegal pointer dereferences, double-free of malloc'd memory, misaligned memory access; integer overflows, divide by zero, memory leaks, concurrent software, deadlock, and data races (i.e. competing writes).
